---
title: "ESM 204 HW 4: Calculating the SCC and Policy Choice under Uncertainty"
author: "Kort Alexander, Eva Wilson, Julia Wilson"
date: "5/19/2022"
output: 
  html_document:
    theme: spacelab
    code_folding: hide
---

```{r setup, include = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(here)
```

```{r}
damages <- read_csv(here("damages.csv"))
warming <- read_csv(here("warming.csv"))
```

### 1. Estimate a quadratic damage function 

```{r}
damages <- damages %>% 
  mutate(warming_sq = warming^2)
damages_model <- lm(data = damages, damages ~ 0 + warming + warming_sq)

warming_range <- damages$warming
damage_predict <- predict(damages_model, list(warming = warming_range, warming_sq = warming_range^2))
damage_predictions <- tibble(warming_range, damage_predict)

ggplot() +
  geom_point(data = damages, aes(x = warming, y = damages)) +
  geom_line(data = damage_predictions, aes(x = warming_range, y = damage_predict), color = "blue") +
  labs(x = "Warming (degrees C)", y = "Damages ($)", title = "Estimated damages (USD) per unit of additional warming (Celsius)") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(axis.title = element_text(face = "bold")) +
  theme(strip.background = element_rect(color= "white", fill = "gray"))
```

***

### 2. Predict damages under the baseline climate and the pulse scenario

```{r}
damage_function <- function(warming) {
  damages <- damages_model$coefficients[1] * warming + damages_model$coefficients[2] * (warming^2)
  return(damages)
}
```

```{r}
warming <- warming %>% 
  mutate(damages_baseline = damage_function(warming_baseline), damages_pulse = damage_function(warming_pulse)) %>% 
  mutate(difference = damage_function(warming_pulse) - damage_function(warming_baseline)) %>% 
  mutate(difference_per_ton = (damage_function(warming_pulse) - damage_function(warming_baseline)) / 35000000000 )

ggplot(warming, aes(x = year, y = damages_baseline)) +
  geom_line(size = 1) +
  labs(x = "Year", y = "Damages ($)", title = "Baseline damages ($)") +
  theme_minimal(14) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(axis.title = element_text(face = "bold")) +
  theme(strip.background = element_rect(color= "white", fill = "gray"))

ggplot(warming, aes(x = year, y = damages_pulse)) +
  geom_line(size = 1) +
  labs(x = "Year", y = "Damages ($)", title = "Damages ($) with pulse") +
  theme_minimal(14) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(axis.title = element_text(face = "bold")) +
  theme(strip.background = element_rect(color= "white", fill = "gray"))

ggplot(warming, aes(x = year, y = difference)) +
  geom_line(size = 1) +
  labs(x = "Year", y = "Damages ($)", title = "Difference in damages ($) with pulse") +
  theme_minimal(14) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(axis.title = element_text(face = "bold")) +
  theme(strip.background = element_rect(color= "white", fill = "gray"))

ggplot(warming, aes(x = year, y = difference_per_ton)) +
  geom_line(size = 1) +
  labs(x = "Year", y = "Damages ($/tonCO2)", title = "Difference in damages with pulse per ton of CO2") +
  theme_minimal(14) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  theme(axis.title = element_text(face = "bold")) +
  theme(strip.background = element_rect(color= "white", fill = "gray"))
```

***

### 3. Plot the SCC against reasonable discount rates

```{r}
SCC_function <- function(r){
  sum <- 0
  for(i in seq(1, 79, 1)){
    sum <- sum + (warming[i, 8] / (1 + r)^i)
  }
  return(sum)
}

discount_rate <- seq(0, 0.1, 0.0025)
SCC <- map(discount_rate, SCC_function) %>% unlist()
SCC_table <- tibble(discount_rate, SCC)

ggplot(SCC_table, aes(x = discount_rate * 100, y = SCC)) +
  geom_line(size = 1) +
  labs(x = "Discount rate (%)", y = "SCC ($/ton)", title = "Social cost of carbon per discount rate") +
  theme_minimal(14) + 
  theme(axis.title = element_text(face = "bold")) +
  theme(strip.background = element_rect(color= "white", fill = "gray")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) 
```

**Interpretation:** Higher discount rates result in a lower SCC; if future climate damages are discounted at a high rate, we would be placing less value on avoiding those damages today.

***

### 4. SCC calculation based on Ramsey Rule

```{r include = FALSE}
r_ramsey <- 0.001 + (2 * 0.01)
SCC_ramsey <- SCC_function(r_ramsey)
r_ramsey
SCC_ramsey
```

```{r}
# plot results
ggplot(SCC_table, aes(x = discount_rate * 100, y = SCC)) +
  geom_line(size = 1) +
  geom_point(aes(x = 0.021 * 100, y = 69.88626), color = "red", size = 4) +
  labs(x = "Discount rate (%)", y = "SCC ($/ton)", title = "Social cost of carbon per discount rate") +
  theme_minimal(14) + 
  annotate("text", x = 3.15, y = 77, label = "Highlighted SCC") +
  theme(axis.title = element_text(face = "bold")) +
  theme(strip.background = element_rect(color= "white", fill = "gray")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) 
```

***

### 5. Consider two possible climate scenarios

```{r include = FALSE}
PV_baseline_damages <- function(r){
  sum <- 0
  for(i in seq(1, 79, 1)){
    sum <- sum + (warming[i, 5] / (1 + r)^i)
  }
  return(sum)
}

damages_polA_warming <- map(warming$warming_baseline * 1.5, damage_function) %>% unlist()
warming$damages_polA_warming <- damages_polA_warming

PV_polA_damages <- function(r){
  sum <- 0
  for(i in seq(1, 79, 1)){
    sum <- sum + (warming[i, 9] / (1 + r)^i)
  }
  return(sum)
}

policy_A_exp_damages <- function(r){
  exp_damages <- 0.5 * (PV_baseline_damages(r)) + 0.5 * (PV_polA_damages(r))
  return(exp_damages)
}

cost_policy_A <- policy_A_exp_damages(0.02) %>% unlist()
cost_policy_A

warming <- warming %>% 
  mutate(warming_polB = ifelse(year <= 2050, warming_baseline, 1.29)) %>% 
  mutate(damages_polB = damage_function(warming_polB))

PV_polB_damages <- function(r){
  sum <- 0
  for(i in seq(1, 79, 1)){
    sum <- sum + (warming[i, 11] / (1 + r)^i)
  }
  return(sum)
}

cost_policy_B <- function(x){
  cost <- x + PV_polB_damages(0.02)
  return(cost)
}

cost_table <- tibble(imp_cost = seq(0, 5e15, 1e13))
cost_table$polA_cost <- cost_policy_A
polB_cost = map(cost_table$imp_cost, cost_policy_B) %>% unlist()
cost_table$polB_cost <- polB_cost 

cost_table <- cost_table %>% 
  pivot_longer(cols = polA_cost:polB_cost, names_to = "policy", values_to = "cost")
```

```{r}
#plot results
ggplot(cost_table, aes(x = imp_cost, y = cost, color = policy)) +
  geom_line(size = 1) +
  labs(x = "Implementation Cost ($)", y = "Total Cost ($)", color = "Policy", title = "Total policy cost by implementation cost") +
  theme_minimal(14) + 
  theme(axis.title = element_text(face = "bold")) +
  theme(strip.background = element_rect(color= "white", fill = "gray")) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) 
```

**Advice to Congress:** We recommend Policy B if the implementation cost is less than 2.19 quadrillion dollars. If the cost exceeds 2.19 quadrillion dollars, we recommend Policy A.

### 6. Evaluating climate policies under a risk-averse scenario

**Advice to Congress:** If society is risk-averse, Policy B will be more attractive. This is because the expected utility cost of Policy A will increase because of the uncertainty involved. Policy B will be the better option for a greater range of implementation costs.
